\documentclass[12pt]{article}
\input{MATH_1060_Preamble.tex}

\setcounter{page}{83}

\begin{document}
\section*{3.7: The Chain Rule}

\boxenv{Learning Objectives.}{Upon successful completion of Section 3.7, you will be able to\dots
		
	\begin{itemize}[leftmargin=6mm]
		\item Answer conceptual questions involving the chain rule.
		\item Identify inner and outer functions for composite functions, then apply the chain rule.
		\item Find derivatives of functions involving the chain rule by using tables and graphs.
		\item Find derivatives of basic functions using the chain rule.
		\item Use the chain rule multiple times and use the chain rule with the product and quotient rules.
		\item Find higher order derivatives using the chain rule.
		\item Find slopes of curves and equations of tangent lines by using the chain rule.
		\item Solve applications involving the chain rule.
	\end{itemize}
	\vspace{-4mm}
}

\vspace{5mm}

\subsection*{Derivatives of Composite Functions}

The \textbf{chain rule} allows us to take derivatives of composite functions.

\vspace{3mm}

\boxenv{The Chain Rule.}{Let $f$ and $g$ be two functions and let $h(x)=f\big( g(x)\big)$ be their composition. The derivative of $h$ is

$$\frac{d}{dx}\Big( f \big( g(x) \big) \Big) = f'\big( g(x)\big) \cdot g'(x).$$

\vspace{-3mm}}

\vspace{5mm}

In words, when applying the chain rule to a composite function, we take the derivative of the \textit{outside} function, keeping the inside function the same, then multiplying by the derivative of the \textit{inside} function.

\Example Find the derivative of $y=\tan\lp x^2-3x\rp$.

\newpage

\paragraph{Alternative Definition.} Another formulation of the chain rule says that if $y=f(u)$ and $u=g(x)$, then the derivative $\frac{dy}{dx}$ is found as follows.

$$\frac{dy}{dx}=\frac{dy}{du}\cdot\frac{du}{dx}$$

\vspace{3mm}

We will usually use the first form shown, but you should be familiar with this form as well.

\vspace{5mm}

\Example If $f(x)=\cos x$ and $g(x)=x^2-2x$, find the derivative of $h(x)=f\big( g(x)\big)$.

\vspace{32mm}

\Example Find $\disp\frac{dy}{dx}$ for $y=\sqrt{4+3x}$.

\vspace{32mm}

\Example Find the derivative of $y=\sec^2 x$. 

\vspace{32mm}

\Example Find the derivative of $f(x)= e^{x^2+x}-\cos 2x$.

\vspace{32mm}

\newpage

\Example Find the derivative of $y=\disp\frac{e^{\pi x}}{e^x+e^{-x}}.$

\vspace{32mm}

\Example Find $\disp\frac{dy}{dx}$ for $y=\lp 2x^3+x^2\rp^{10}$.

\vspace{32mm}

\Example Find the derivative of $g(x)=x^2\lp x^3+2\rp ^4$.

\vspace{32mm}

\Example Find the derivative of $f(x)=\sqrt{x^4+x^3+x^2+x}$.

\vspace{32mm}

\Example Find $\disp\frac{dy}{dx}$ for $y=\sin\big(\cos\lp \tan x\rp\big)$.

\newpage

\Example Find $f'$ given that $f(x)=x^3 g(2x+1)$, where $g$ is differentiable.

\vspace{40mm}

\Example Find $\disp\frac{dy}{dx}$ for $y=\tan \lp \sin 2x\rp$.

\vspace{40mm}

\Example Differentiate $f(x)=e^{k\tan \sqrt{x}}$ for $k\in\R$, $k\neq 0$.

\vspace{40mm}

\Example Find the derivative of $g(t)=\sin^3(3t)$ and $f(t)=3\sin t^3$.

\newpage

\Example Find all values of $x$ where $y=(3x-1)^2(5-x)^4$ is horizontal.

\vspace{50mm}

\Example Suppose $f$ is differentiable with $f'(0)=3$ and $f'\lp\frac{\sqrt{2}}{2}\rp=2$. \\

If $g(x)=f(\sin x)+f\lp x-\disp\frac{\pi}{4}\rp$, calculate $g'\lp\disp\frac{\pi}{4}\rp$.

\vspace{50mm}

\Example Suppose $f$ and $g$ are two lines with slopes $m$ and $\disp\frac{1}{m}$, respectively, where $m\neq 0$. If $y=f\big(g(x)\big)$, what is the value of $y'(c)$ where $c\in\R$?


\end{document}