\documentclass[12pt]{article}
\input{MATH_1060_Preamble.tex}

\setcounter{page}{72}

\begin{document}
\section*{3.4: The Product and Quotient Rules}

\boxenv{Learning Objectives.}{Upon successful completion of Section 3.4, you will be able to\dots
		
	\begin{itemize}[leftmargin=6mm]
		\item Answer conceptual questions involving the product and quotient rules.
		\item Find derivatives using two different methods.
		\item Find derivatives of products and quotients of functions involving exponentials.
		\item Find derivatives of products and quotients of algebraic expressions.
		\item Find derivatives using the extended power rule.
		\item Find slopes and equations of tangent lines of functions involving products and \\ quotients.
		\item Solve applications involving the product rule and quotient rule.
		\item Find higher order derivatives of products and quotients.
		\item Find derivatives of products and quotient rules using given values or graphs.
	\end{itemize}
	\vspace{-4mm}
}

\vspace{5mm}

\subsection*{The Product Rule}

To differentiate a product of two functions, we use a technique called the \textbf{product rule}.

\vspace{3mm}

\boxenv{The Product Rule.}{For two differentiable functions $f$ and $g$, the derivative of their product $fg$ is

$$\lp fg\rp '=f'g+fg'.$$

\vspace{3mm}

Alternatively, the product rule can be expressed as follows.

\vspace{15mm}}

\vspace{5mm} For example, consider the function $y=x^2e^x$.

$$\frac{d}{dx}\lp x^2 e^x\rp = 2xe^x+x^2e^x$$

\newpage

\textit{Proof of the Product Rule.} Suppose $f$ and $g$ are both differentiable and let $h(x)=f(x)g(x)$.
\begin{align*}
h'(x) &= \lim_{h\to 0}\frac{h(x+h)-h(x)}{h} \\
&= \lim_{h\to 0}\frac{f(x+h)g(x+h)-f(x)g(x)}{h}\\
&= \lim_{h\to 0}\frac{f(x+h)g(x+h) {\color{blue}\overbrace{- f(x)g(x+h) + f(x)g(x+h)}^{=0}} - f(x)g(x)}{h}\\
&= \lim_{h\to 0}\frac{g(x+h)\big(f(x+h) - f(x)\big) + f(x)\big(g(x+h) - g(x)\big)}{h} \\
	      &= \lim_{h\to 0}g(x+h)\underbrace{\lim_{h\to 0}\frac{f(x+h)-f(x)}{h}}_{=f'(x)} + f(x)\underbrace{\lim_{h\to 0}\frac{g(x+h) - g(x)}{h}}_{=g'(x)} \\
	      &= g(x) f'(x) + f(x) g'(x)
\end{align*}
\begin{flushright}
	$\square$
\end{flushright}

\Example Find the derivative of $f(x)=x^4e^x$.

\vspace{50mm}

\Example Find the derivative of $f(x)=2e^x\lp x^3-3\rp$.

\newpage

\Example Determine a formula for the $n^\text{th}$ derivative of $f(x)=xe^x$.

\vspace{60mm}

\Example Find the derivative of $g(x)=6x^2\lp 3+2x^{-3}\rp$.

\vspace{50mm}

\Example Find the derivative of $h(x)=\disp\frac{70+2x^2}{x^3}$.

\newpage

\subsection*{The Quotient Rule}

\boxenv{The Quotient Rule.}{For two differentiable functions $f$ and $g$, where $g(x)\neq 0$, the derivative of their quotient $\frac{f}{g}$ is

$$\lp\frac{f}{g}\rp '=\frac{f'g-fg'}{g^2}.$$

\vspace{3mm}

Alternatively, the quotient rule can be expressed as follows.

\vspace{15mm}}

\vspace{3mm}

Note that subtraction is not commutative, i.e.\ $x-y\neq y-x$, so the order of subtraction in the numerator is critical.

\vspace{10mm}

\Example Evaluate the expression $\disp\frac{d}{dx}\lp\frac{e^x}{x+1}\rp$.

\vspace{50mm}

\Example Find the derivative of $y=\disp\frac{x+2}{3x^2-x}$.

\newpage

\Example Find the derivative of $g(x)=\disp\frac{2+xe^x}{x^2-2}$.

\vspace{55mm}

\Example Find the equation of the line tangent to $y=\disp\frac{2x}{x+1}$ at $x=2$.

\vspace{55mm}

\Example if $h(x)=f(x)g(x)$, find $h'(2)$ given the following information.

$$f(2)=-3\hspace{15mm} f'(2)=-2 \hspace{15mm} g(2)=4 \hspace{15mm} g'(2)=7$$

\newpage 

\Example Suppose $f$ is a differentiable function. Find an expression for the derivative of each of the following functions.

\begin{enumerate}
\item[\tc{1}] $y=x^2f(x)$

\vspace{35mm}

\item[\tc{2}] $y=\disp\frac{\sqrt{x}+xf(x)}{e^x}$

\vspace{35mm}
\end{enumerate}

\Example Find the derivative for each of the following functions.

\begin{enumerate}
\item[\tc{1}] $g(x)=\disp\frac{2x^5+x-1}{e^x+9}$

\vspace{35mm}

\item[\tc{2}] $f(x)=\sqrt{x}\lp 3x^7-2\rp$
\end{enumerate}
\end{document}