\documentclass[12pt]{article}
\input{MATH_1080_Preamble.tex}

\setcounter{page}{116}

\begin{document}
\section*{10.5: Comparison Tests}

\boxenv{Learning Objectives.}{Upon successful completion of Section 10.5, you will be able to\dots
		
	\begin{itemize}[leftmargin=6mm]
		\item Answer conceptual questions involving the Comparison Tests.
		\item Use the Comparison Test or Limit Comparison Tests to determine series convergence or divergence.
		\item Determine series convergence or divergence using a test of your choice.
	\end{itemize}
	\vspace{-4mm}
}

\vspace{5mm}

\subsection*{The (Direct) Comparison Test}

Similar to the Comparison Theorem for Improper Integrals that was introduced in Section~8.9, we can compare a given infinite series to a series that we know to be convergent or divergent. We will usually do comparisons to a $p$-series or a geometric series.

\vspace{5mm}

\boxenv{Comparison Test.}{Suppose that $\sum a_n$ and $\sum b_n$ are series with positive terms. Then\dots
\vspace{2mm}
\begin{itemize}
\item[(1)] If $\sum b_n$ converges and $a_n\leq b_n$ for all $n$, then $\sum a_n$ converges.
\vspace{2mm}
\item[(2)] If $\sum b_n$ diverges and $a_n\geq b_n$ for all $n$, then $\sum a_n$ diverges.
\end{itemize}
}

\Example $\disp\sum_{n=1}^\infty\dfrac{n}{2n^3+1}$

\newpage

\Example $\disp\sum_{k=2}^\infty \dfrac{\ln k}{k}$

\vfill

\subsubsection*{Important Notes on Using Comparison Tests}

\vspace{60mm}

\newpage

\subsection*{The Limit Comparison Test}

\boxenv{Limit Comparison Test.}{Suppose that $\sum a_n$ and $\sum b_n$ are series with positive terms.\\

If $\disp\lim_{n\to\infty}\dfrac{a_n}{b_n}=c$, where $0<c<\infty$,\\

then either $\sum a_n$ and $\sum b_n$ both converge or both diverge.}

\Example $\disp\sum_{n=1}^\infty\dfrac{n}{2n^3-1}$

\newpage

\Example $\disp\sum_{n=1}^\infty\dfrac{\sqrt{n^4+1}}{n^3+n^2}$

\vfill

\Example $\disp\sum_{n=1}^\infty\tan\left(\dfrac{1}{n}\right)$

\vfill

\newpage

\Example $\disp\sum_{n=1}^\infty\dfrac{n+4^n}{n+6^n}$

\end{document}